---
title: "Analysis of Bike Sharing Dataset"
author: "Ayano Yamamoto"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up
```{r}
# Create a list of required packages
needed_packages <- c("tidyverse", "cowplot", "ggpubr", "semTools", "FSA", "e1071", "psych", "car", "effectsize", "gmodels", "sjstats")

# Extract not installed packages
not_installed <- needed_packages[!(needed_packages %in% installed.packages()[ , "Package"])]    

# Install not installed packages
if(length(not_installed)) install.packages(not_installed) 

# Load the required packages
library(tidyverse)
library(cowplot)
library(ggpubr)
library(semTools)
library(FSA)
library(e1071)
library(psych)
library(car)
library(effectsize)
library(gmodels)
library(sjstats)

# Importing the the file `BikeSharing (By Day).csv` with `instant` variable as row names.
bike_sharing <- read.csv(file = 'Bike Sharing/BikeSharing (By Day).csv', 
                         row.names = 'instant', 
                         stringsAsFactors = FALSE)

# Check the first few rows
head(bike_sharing)

# Display an overview of the dataset
str(bike_sharing)

# Check for duplicated rows
sum((duplicated(bike_sharing)))

# Check for NAs
colSums(is.na(bike_sharing))
```

# Question 1
#### Does the temperature (actual) impact the total number of bikes hired per day?

## Description of `cnt`
```{r}
# Datatype
typeof(bike_sharing$cnt)

# Sample size
sum(!is.na(bike_sharing$cnt))

# Overview
summary(bike_sharing$cnt)

# Standard deviation
round(sd(bike_sharing$cnt), 3)

# Create a histogram with density curve and a box plot of total number of bikes hired per day
cnt1 <- 
  bike_sharing %>%
    ggplot(aes(x = cnt)) +
    geom_histogram(aes(y = ..density..),binwidth = 250, fill = "#C6CDF7") +
    geom_density(lwd = 1.2, linetype = 2, colour = "#E6A0C4", fill = "#E6A0C4", alpha = 0.2) +
    # draw a dashed line at mean
    geom_vline(aes(xintercept = mean(cnt), colour = "mean"), linetype = "dashed") + 
    # draw a red line at median
    geom_vline(aes(xintercept = median(cnt), colour = "median")) + 
    scale_color_manual(name = "Central Tendency", values = c(median = "#C93312", mean = "black")) +
    labs(x = "", y = "Frequency") +
    ggtitle("Figure 1: Distribution of total number of bikes hired per day")

cnt2 <-
    bike_sharing %>%
    ggplot(aes(x = cnt)) +
    geom_boxplot(fill = "#D8A499", alpha = 0.5) +
    labs(x = "Total number of bikes hired per day", y = "") +
    theme(axis.text.y = element_blank(), 
          axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank())

plot_grid(cnt1, cnt2, ncol = 1, align = "v", axis = "lr", rel_heights = c(4, 1))

# Create a quantile-quantile plot of total number of bikes hired per day
bike_sharing %>%
  ggqqplot(x = "cnt", color = "#7294D4") +
  ggtitle("Figure 2: Quantile-quantile plot of total number of bikes hired per day")
    
# Quantify skew
cnt_skew <- semTools::skew(bike_sharing$cnt)
cnt_kurt <- semTools::kurtosis(bike_sharing$cnt)
cnt_skew[1]/cnt_skew[2]

# Quantify excess kurtosis
cnt_kurt[1]/cnt_kurt[2]

# Share of Z score outside +- 3.29
cnt_zscore <- abs(scale(bike_sharing$cnt))
FSA::perc(as.numeric(cnt_zscore), 3.29, "gt")
```
```
Number of bikes hired per day can be considered to follow a normal distribution (m = `r round(mean(bike_sharing$cnt, na.rm=TRUE),2)`, sd = `r round(sd(bike_sharing$cnt, na.rm=TRUE),2)`, n = `r length(bike_sharing$cnt)-sum(is.na(bike_sharing$cnt))`, 100% of standardized scores falling within +/- 3.29).
```

## Description of `temp`
```{r}
# Datatype
typeof(bike_sharing$temp)

# Sample size
sum(!is.na(bike_sharing$temp))

# Overview
summary(bike_sharing$temp)

# Interquartile range
round(sd(bike_sharing$temp), 3)

# Create a histogram with density curve and a box plot of total number of bikes hired per day
cnt1 <- 
  bike_sharing %>%
    ggplot(aes(x = temp)) +
    geom_histogram(aes(y = ..density..),binwidth = 0.025, fill = "#C6CDF7") +
    geom_density(lwd = 1.2, linetype = 2, colour = "#E6A0C4", fill = "#E6A0C4", alpha = 0.2) +
    # draw a dashed line at mean
    geom_vline(aes(xintercept = mean(temp), colour = "mean"), linetype = "dashed") + 
    # draw a red line at median
    geom_vline(aes(xintercept = median(temp), colour = "median")) + 
    scale_color_manual(name = "Central Tendency", values = c(median = "#C93312", mean = "black")) +
    labs(x = "", y = "Frequency") +
    ggtitle("Figure 3: Distribution of normalized feeling temperature in Celsius")

cnt2 <-
    bike_sharing %>%
    ggplot(aes(x = temp)) +
    geom_boxplot(fill = "#D8A499", alpha = 0.5) +
    labs(x = "Normalized feeling temperature in Celsius", y = "") +
    theme(axis.text.y = element_blank(), 
          axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank())

plot_grid(cnt1, cnt2, ncol = 1, align = "v", axis = "lr", rel_heights = c(4, 1))

# Create a quantile-quantile plot of total number of bikes hired per day
bike_sharing %>%
  ggqqplot(x = "temp", color = "#7294D4") +
  ggtitle("Figure 4: Quantile-quantile plot of normalized temperature in Celsius")

# Quantify skew
temp_skew <- semTools::skew(bike_sharing$temp)
temp_kurt <- semTools::kurtosis(bike_sharing$temp)
temp_skew[1]/temp_skew[2]

# Quantify excess kurtosis
temp_kurt[1]/temp_kurt[2]

# Share of Z score outside +- 3.29
temp_zscore <- abs(scale(bike_sharing$temp))
FSA::perc(as.numeric(temp_zscore), 3.29, "gt")
```
```
Temperature (actual) can be considered to follow a normal distribution (m = `r round(mean(bike_sharing$temp, na.rm=TRUE),2)`, sd = `r round(sd(bike_sharing$temp, na.rm=TRUE),2)`, n = `r length(bike_sharing$temp)-sum(is.na(bike_sharing$temp))`, 100% of standardized scores falling within +/- 3.29). Therefore we can use a parametric correlation test.
```

## Correlation between `cnt` and `temp`
```{r}
# Create a scatter plot with a regression line
ggplot(bike_sharing, aes(x = temp, y = cnt)) +
    geom_point(color = "#7294D4") +
    geom_smooth(method = "lm",
                color = "#C93312",
                size = 1.5,
                se = FALSE) +
    labs(x = "Normalized feeling temperature in Celsius",
         y = "Total number of bikes hired per day")+
    ggtitle("Figure 5: Plot of temperature versus total number of bikes hired per day")

# Pearson correlation
cor.test(bike_sharing$cnt, bike_sharing$temp, method = "pearson")
```
```
The relationship between the total number of bikes hired per day (count of total rental bikes including both casual and registered) and temperature (actual) (normalised temperature in Celsius) was investigated using a Pearson correlation. A statistically significant result was found indicating a strong positive correlation (r  = .63, n = 73, p < .001).
```

# Question 2
#### Does the level of humidity impact the total number of bikes hired per day?

## Description of `hum`
```{r}
# Datatype
typeof(bike_sharing$hum)

# Sample size
sum(!is.na(bike_sharing$hum))

# Overview
summary(bike_sharing$hum)

# Interquartile range
round(sd(bike_sharing$hum), 2)

# Display observations of hum = 0
bike_sharing %>%
  filter(hum == 0)

# Create a histogram with density curve and a box plot of total number of bikes hired per day
cnt1 <- 
  bike_sharing %>%
    ggplot(aes(x = hum)) +
    geom_histogram(aes(y = ..density..),binwidth = 0.025, fill = "#C6CDF7") +
    geom_density(lwd = 1.2, linetype = 2, colour = "#E6A0C4", fill = "#E6A0C4", alpha = 0.2) +
    # draw a dashed line at mean
    geom_vline(aes(xintercept = mean(hum), colour = "mean"), linetype = "dashed") + 
    # draw a red line at median
    geom_vline(aes(xintercept = median(hum), colour = "median")) + 
    scale_color_manual(name = "Central Tendency", values = c(median = "#C93312", mean = "black")) +
    labs(x = "", y = "Frequency") +
    ggtitle("Figure 6: Distribution of normalized humidity")

cnt2 <-
    bike_sharing %>%
    ggplot(aes(x = hum)) +
    geom_boxplot(fill = "#D8A499", alpha = 0.5) +
    labs(x = "Normalized humidity", y = "") +
    theme(axis.text.y = element_blank(), 
          axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank())

plot_grid(cnt1, cnt2, ncol = 1, align = "v",  axis = "lr", rel_heights = c(4, 1))

# Create a quantile-quantile plot of total number of bikes hired per day
bike_sharing %>%
  ggqqplot(x = "hum", color = "#7294D4") +
  ggtitle("Figure 7: Quantile-quantile plot of normalized humidity")

# Quantify skew
hum_skew <- semTools::skew(bike_sharing$hum)
hum_kurt <- semTools::kurtosis(bike_sharing$hum)
hum_skew[1]/hum_skew[2]

# Quantify excess kurtosis
hum_kurt[1]/hum_kurt[2]

# Share of Z score outside +- 3.29
hum_zscore <- abs(scale(bike_sharing$hum))
FSA::perc(as.numeric(hum_zscore), 3.29, "gt")
```
```
The level of humidity can be considered to follow a normal distribution (m = `r round(mean(bike_sharing$hum, na.rm=TRUE),2)`, sd = `r round(sd(bike_sharing$hum, na.rm=TRUE),2)`, n = `r length(bike_sharing$hum)-sum(is.na(bike_sharing$hum))`, 99.86% of standardized scores falling within +/- 3.29). Therefore we can use a parametric correlation test.
```

## Correlation between `cnt` and `hum`
```{r}
# Create a scatter plot with a regression line
ggplot(bike_sharing, aes(x = hum, y = cnt)) +
    geom_point(color = "#7294D4") +
    geom_smooth(method = "lm",
                color = "#C93312",
                size = 1.5,
                se = FALSE) +
    labs(x = "Normalized humidity",
         y = "Total number of bikes hired per day")+
    ggtitle("Figure 8: Plot of normalized humidity versus total number of bikes hired per day")

# Pearson correlation
cor.test(bike_sharing$cnt, bike_sharing$hum, method = "pearson")
```
```
The relationship between the total number of bikes hired per day (count of total rental bikes including both casual and registered) and the level of humidity (normalised humidity) was investigated using a Pearson correlation. A statistically significant result was found indicating a weak negative correlation (r  = -.10, n = 729, p = .01). 
```

# Question 3
#### Does the total number of bikes hired per day vary according to whether a day is a regular weekday or a weekend?

## Description of `workingday`
```{r}
# Datatype
typeof(bike_sharing$workingday)

# Sample size
sum(!is.na(bike_sharing$workingday))

# Set of all possible values
unique(bike_sharing$workingday)

# Frequency and percentages
bike_sharing %>%
  group_by(workingday) %>%
  summarise(n = n(), percentage = round(n() / nrow(bike_sharing) * 100, 2))

# Create a bar chart of whether a day is a regular weekday or a weekend
bike_sharing %>%
  ggplot(aes(x = as.factor(workingday), fill = as.factor(workingday))) +
  geom_bar() +
  scale_fill_manual(name = "Workingday", values = c("#E6A0C4", "#C6CDF7")) +
  labs(x = "Workingday", y = "Frequency") +
  ggtitle("Figure 9: Distribution of  whether a day is a regular weekday or a weekend")
```

## Comparison between `cnt` and `workingday`
```{r}
# Descriptive statistics by group
psych::describeBy(bike_sharing$cnt, bike_sharing$workingday, mat=TRUE)

# Conduct Levene's test for homogeneity of variance
car::leveneTest(cnt ~ as.factor(workingday), data = bike_sharing)

# Conduct the t-test using the var.equal = FALSE option
stats::t.test(cnt ~ as.factor(workingday), var.equal = FALSE, data = bike_sharing)

# Calculate Cohen's d arithmetically
workingday_ttest <- stats::t.test(cnt ~ as.factor(workingday), var.equal = FALSE, data = bike_sharing)
workingday_effcd <- round((2*workingday_ttest$statistic)/sqrt(workingday_ttest$parameter),2)
effectsize::t_to_d(t = workingday_ttest$statistic, workingday_ttest$parameter)
```
```
An independent-samples t-test was conducted to compare the total number of bikes hired per day for regular weekdays and weekends. No statistically significant difference in the total number of bikes hired per day was found (M = `r round(mean(bike_sharing$cnt[bike_sharing$workingday == 0], na.rm=TRUE),2)`, SD =`r round(sd(bike_sharing$cnt[bike_sharing$workingday == 0], na.rm = TRUE), 2)` for weekends, M = `r round(mean(bike_sharing$cnt[bike_sharing$workingday == 1], na.rm=TRUE),2)`, SD=`r round(mean(bike_sharing$cnt[bike_sharing$workingday == 1], na.rm = TRUE), 2)` for weekdays), (t(`r workingday_ttest$parameter`)= `r round(workingday_ttest$statistic, 3)`, p = 0.1101. Cohen's d also indicated a small effect size (`r workingday_effcd`).
```

# Question 4
#### Does the total number of bikes hired per day vary by the day of the week?

## Description of `weekday`
```{r}
# Datatype
typeof(bike_sharing$weekday)

# Sample size
sum(!is.na(bike_sharing$weekday))

# Set of all possible values
unique(bike_sharing$weekday)

# Frequency and percentages
bike_sharing %>%
  group_by(weekday) %>%
  summarise(n = n(), percentage = round(n() / nrow(bike_sharing) * 100, 2))

# Create a bar chart of days of the week
bike_sharing %>%
  ggplot(aes(x = as.factor(weekday), fill = as.factor(weekday))) +
  geom_bar() +
  scale_fill_brewer(name = "Day of the Week", palette="BuPu") +
  labs(x = "Day of the Week", y = "Frequency") +
  ggtitle("Figure 10: Distribution of days of the week")
```

## Comparison between `cnt` and `weekday`
```{r}
# Descriptive statistics by group
psych::describeBy(bike_sharing$cnt, bike_sharing$weekday, mat=TRUE)

# Store the output to use in our final reporting of the outcomes of ANOVA
weekday_descrip <- psych::describeBy(bike_sharing$cnt, bike_sharing$weekday, mat=TRUE)

# Conduct Bartlett’s test for homogeneity of variance
stats::bartlett.test(cnt ~ as.factor(weekday), data = bike_sharing)

# Compute the analysis of variance using the var.equal = TRUE option
weekday_anova <- stats::oneway.test(cnt ~ as.factor(weekday), data = bike_sharing, var.equal = TRUE)

# Summary of the analysis
weekday_anova

# Compute Eta squared
weekday_effes <- effectsize::effectsize(weekday_anova)
weekday_effes

# Store the relevant pieces of the output from ANOVA in variables to use for reporting
# Degrees of freedom
weekday_df1 <- weekday_anova$parameter[1]
weekday_df2 <- weekday_anova$parameter[2]

# F statistic
weekday_Fstat <- round(weekday_anova$statistic, 3)

# P-value
weekday_pval <- round(weekday_anova$p.value,2)
```
```
A one way ANOVA test was conducted to compare the total number of bikes hired per day and days of the week. No statistically significant difference in the scores for the total number of bikes hired per day and days of the week was found  (F(6, `r weekday_df1`, `r weekday_df2`) = `r round(weekday_anova$statistic,2)`, p=`r weekday_pval`), (
(M = `r round(weekday_descrip$mean[1],2)`, SD = `r round(weekday_descrip$sd[1],2)`) for 0, 
(M = `r round(weekday_descrip$mean[2],2)`, SD = `r round(weekday_descrip$sd[2],2)`) for 1,  
(M = `r round(weekday_descrip$mean[3],2)`, SD = `r round(weekday_descrip$sd[3],2)`) for 2,
(M = `r round(weekday_descrip$mean[4],2)`, SD = `r round(weekday_descrip$sd[4],2)`) for 3,
(M = `r round(weekday_descrip$mean[5],2)`, SD = `r round(weekday_descrip$sd[5],2)`) for 4,
(M = `r round(weekday_descrip$mean[6],2)`, SD = `r round(weekday_descrip$sd[6],2)`) for 5,
(M = `r round(weekday_descrip$mean[7],2)`, SD = `r round(weekday_descrip$sd[7],2)`) for 6).
A small effect size was also indicated by the eta squared value (`r round(weekday_effes$Eta2,2)`).

```

# Question 5
#### Is the weather situation related to the season?

## Description of `weathersit`
```{r}
# Datatype
typeof(bike_sharing$weathersit)

# Sample size
sum(!is.na(bike_sharing$weathersit))

# Set of all possible values
unique(bike_sharing$weathersit)

# Frequency and percentages
bike_sharing %>%
  group_by(weathersit) %>%
  summarise(n = n(), percentage = round(n() / nrow(bike_sharing) * 100, 2))

# Create a bar chart of weather situation
bike_sharing %>%
  ggplot(aes(x = as.factor(weathersit), fill = as.factor(weathersit))) +
  geom_bar() +
  scale_fill_manual(name = "Weathersit", values = c("#E6A0C4", "#C6CDF7", "#7294D4")) +
  labs(x = "Weather situation", y = "Frequency") +
  ggtitle("Figure 11: Distribution of weather situation")
```

## Description of `season`
```{r}
# Datatype
typeof(bike_sharing$season)

# Sample size
sum(!is.na(bike_sharing$season))

# Set of all possible values
unique(bike_sharing$season)

# Frequency and percentages
bike_sharing %>%
  group_by(season) %>%
  summarise(n = n(), percentage = round(n() / nrow(bike_sharing) * 100, 2))

# Create a bar chart of season
bike_sharing %>%
  ggplot(aes(x = as.factor(season), fill = as.factor(season))) +
  geom_bar() +
  scale_fill_manual(name = "season", values = c("#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4")) +
  labs(x = "Season", y = "Frequency") +
  ggtitle("Figure 12: Distribution of season")
```

## Comparison between `weathersit` and `season`
```{r}
# Conduct Chi-square test
gmodels::CrossTable(bike_sharing$season, bike_sharing$weathersit, chisq = TRUE, expected = TRUE, sresid = TRUE, format = "SPSS")

# Calculate effect size
weathersit_season <- xtabs(~ weathersit + season, data = bike_sharing)
sjstats::cramer(weathersit_season)
```
```
A Chi-square test for independence indicated a statistically significant association between the weather situations and the seasons, 2(6, n = 731) = 14.884, p = 0.0212, V = 0.101). 
````